Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering

### 2014-09-12 evaluation.tex
Revised Evaluation.

In order to explore the generality of the architecture for
different monitors, we implemented four different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), and instruction-mix profiling (IMP).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations. Additionally, recent
work has shown that this information can be used to detect malicious software \cite{tang-raid14}.

@@

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, DIFT, and IMP.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 5x to 16x with an average
of 10x. BC shows normalized execution times of 8-24x with an average of
16x while DIFT shows normalized execution times of 4x-14x with an average of
9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
These monitors show high overheads because they require several
instructions to process each monitored instruction from the main core. Even
small increases in the number of instructions needed by the monitor lead to
high overheads when multiplied across the main program's execution. For
example, because of the large metadata requirements of UMC, BC, and DIFT, these
monitors must dynamically allocate memory for new metadata. On the other hand,
IMP does not require this dynamic allocation and is overall a simple monitor.
Thus, its overheads are much lower.

@@

\subsection{Filtering of Monitoring Events}

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overheads with filtering null metadata for UMC, BC, and DIFT.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 1.7-7x with an average of 4x with null metadata filtering.
This is an average reduction in overheads of 65\%. For BC, normalized execution
times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
times for filtered BC is 1.1-8.3x. This is not surprising since in the baseline
implementation all loads and stores needed to be monitored. However, with
filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
on average after filtering, an 88\% reduction in monitoring overheads, with a
maximum overhead of 31\% after filtering. This is due, in part, to the fact
that
for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering. IMP does not have
null metadata and thus is not affected by null metadata filtering.

@@

% BC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for array bounds check.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure*}

% UMC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

We see that by varying the overhead budget, the coverage achieved also varies.
With only a 10\% overhead budget, array bounds check still
achieves 82\% monitoring coverage on average. The high coverage
achieved with such low overheads is due to two main effects.  The first is that
monitoring can be done in parallel, providing monitoring coverage without
introducing overheads. The second effect is that, although we aggressively
filter out monitoring events, there may still exist a large number of
monitoring events that do not lead to security or reliability checks. As a
result, dropping these events can reduce overheads without a large impact on
monitoring coverage.

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 10\% overhead budget, UMC achieves 30\% monitoring coverage on average.
By increasing this overhead budget to 50\%, UMC achieves 48\% coverage. Even
higher coverage can be achieved by allowing higher overheads.

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We compare these percentages against the case when full monitoring
is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the max error across
instruction types for each benchmark and for varying overheads. We see that
with a 10\% overhead a max error of 9\% is observed and an average of less than
4\%. This error decreases as overhead is increased.

% Flex UMC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
    \label{fig:evaluation.fpga_umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

% UMC exec time
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
    \vspace{-0.2in}
    \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_exec_time}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC coverage across policies
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_coverage}
    \vspace{-0.1in}
  \end{center}
\end{figure}
In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of independent monitoring flows.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that source-dropping
consistently achieves higher coverage than unrestricted dropping. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

% BC exec time
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
    \vspace{-0.2in}
    \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
    \label{fig:evaluation.bc_exec_time}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% BC coverage across policies
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage for array bounds check for different dropping policies.}
    \label{fig:evaluation.bc_coverage}
    \vspace{-0.2in}
  \end{center}
\end{figure}

Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
benchmarks, source-only drop fails to meet the specified overhead target.
The overshoot of the overhead target is quite high with an overhead difference
of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
Since only array allocations are considered source events for BC, it is
difficult for source-only dropping to match overhead targets. Although
Figure~\ref{fig:evaluation.bc_coverage} again shows higher coverage for
source-only dropping, this is in part due to the fact that is running with
higher overheads.

@@

% Area and Power Overheads
\begin{table}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/area_power}
    \end{footnotesize}
    \caption{Average power overhead for dropping hardware at a 50\% overhead
    budget. Percentages in parentheses are normalized to the main core
    power.}
    \vspace{-0.2in}
    \label{tab:evaluation.area_power}
  \end{center}
\end{table}

@@

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 50\% monitoring overhead target. The peak power is 41-72 mW, which is
less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43
mW, corresponding to 5-11\% of the main core's runtime power.

>>>>

In order to explore the generality of the architecture for
different monitors, we implemented four different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), and instruction-mix profiling (IMP).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations. Additionally, recent
work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
Note that, although our dataflow engine performs similar operations to the DIFT
monitor, the dataflow engine operates on smaller, 2-bit flags.

@@

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 5x to 16x with an average
of 10x. BC shows normalized execution times of 8-24x with an average of
16x while DIFT shows normalized execution times of 4x-14x with an average of
9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
% These monitors show high overheads because they require several
% instructions to process each monitored instruction from the main core. Even
% small increases in the number of instructions needed by the monitor lead to
% high overheads when multiplied across the main program's execution.

@@

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overheads with filtering null metadata.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 1.7-7x with an average of 4x with null metadata filtering.
This is an average reduction in overheads of 65\%. For BC, normalized execution
times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
times for filtered BC is 1.1-8.3x. 
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
on average after filtering, an 88\% reduction in monitoring overheads, with a
maximum overhead of 31\% after filtering. This is due, in part, to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering. IMP does not have
null metadata and thus is not affected by null metadata filtering.
These overheads are similar to previous results for multi-core monitoring with
minimal hardware costs such as LBA \cite{lba-asid06} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for array bounds check.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

We see that by varying the overhead budget, the coverage achieved also varies.
With only a 1.1x overhead target, array bounds check still
achieves 84\% monitoring coverage on average. The high coverage
achieved with such low overheads is due to two main effects.  The first is that
monitoring can be done in parallel, providing monitoring coverage without
introducing overheads. The second effect is that, although we aggressively
filter out monitoring events, there may still exist a large number of
monitoring events that do not lead to security or reliability checks. As a
result, dropping these events can reduce overheads without a large impact on
monitoring coverage.

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We compare these percentages against the case when full monitoring
is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the max error across
instruction types for each benchmark and for varying overheads. We see that
with a 1.1x overhead a max error of 9\% is observed and an average of less than
4\%. This error decreases as overhead is increased and is 0\% for all
benchmarks once a 6x overhead target is used.

\subsection{FPGA-based Monitor}
\label{sec:evaluation.fpga}

% Flex UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
    \label{fig:evaluation.fpga_umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% % UMC exec time
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
%     \vspace{-0.2in}
%     \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
%     \label{fig:evaluation.umc_exec_time}
%     \vspace{-0.2in}
%   \end{center}
% \end{figure}
% 
% % UMC coverage across policies
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
%     \vspace{-0.2in}
%     \caption{Coverage for uninitialized memory check for different dropping policies.}
%     \label{fig:evaluation.umc_coverage}
%     \vspace{-0.1in}
%   \end{center}
% \end{figure}

\begin{figure}
  \begin{center}
  \vspace{-0.2in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \vspace{-0.2in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \caption{Comparing dropping policies for UMC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.umc_policies}
\end{figure}

In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
The results are shown for an overhead target of 1.5x.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of independent monitoring flows.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping. On average, unrestricted
dropping achieves 46\% coverage while source-only dropping achieves 49\%
coverage. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

% % BC exec time
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
%     \vspace{-0.2in}
%     \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
%     \label{fig:evaluation.bc_exec_time}
%     \vspace{-0.2in}
%   \end{center}
% \end{figure}
% 
% % BC coverage across policies
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
%     \vspace{-0.2in}
%     \caption{Coverage for array bounds check for different dropping policies.}
%     \label{fig:evaluation.bc_coverage}
%     \vspace{-0.2in}
%   \end{center}
% \end{figure}

\begin{figure}
  \begin{center}
  \vspace{-0.2in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \vspace{-0.2in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \caption{Comparing dropping policies for BC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.bc_policies}
\end{figure}

Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
benchmarks, source-only drop fails to meet the specified overhead target.
The overshoot of the overhead target is quite high with an overhead differences
of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
Since only array allocations are considered source events for BC, it is
difficult for source-only dropping to match overhead targets. Although
Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
source-only dropping, this is largely due to the fact that is running with
higher overheads.

@@

% Area and Power Overheads
\begin{table}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/area_power}
    \end{footnotesize}
    \caption{Average power overhead for dropping hardware. Percentages 
    are normalized to the main core power.}
    \vspace{-0.2in}
    \label{tab:evaluation.area_power}
  \end{center}
\end{table}

@@

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 1.5x monitoring overhead target. The peak power is 41-72 mW, which is
less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43
mW, corresponding to 5-11\% of the main core's runtime power.
